---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# innermap

<!-- badges: start -->
<!-- badges: end -->

The `innermap` package presents the `innermap` family functions. These in turn, allow the use of map family function but using elements of the same vectors. This is not the same thing of `purrr::accumulate` function, since it does not uses any output made by the function in the process. Let's see an example:

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("MarceloRTonon/innermap")
```


thus I need to do a matrice of R and demand that I use list to hold multiples matrices

## Example


```{r}
library(innermap)

```



```{r}
inputVec <- c(1:30)

inputVec
```

Now suppose you want to sum every element of the `inputVec` vector with the following one. Using `base r` you would have to do:


```{r}
outputVec <- vector(mode = "double",
                    length = (length(inputVec)-1)
                    )

for(i in seq_along(inputVec)){
  outputVec[i] <- inputVec[i] + inputVec[i+1]
}

outputVec
```

As we know, using `for` in R is neither very optimized or a good practice. It is possible to do this using the `purrr` package:

```{r}
library(purrr)

outputVec <- purrr::map2_dbl(c(1:(length(inputVec)-1)),
                     c(2:length(inputVec)),
                     function(x,y) inputVec[x] + inputVec[y])

outputVec
```

But one can easily argue that writing explicitly the lengths of the vector in the arguments is at least very anoying, when very confusing to read. The `innermap` approach do the same thing that was done in the example prior, but in the backstage. Let's use `innermap_dbl` in this example:


```{r}

outputVec <- innermap::innermap_dbl(inputVec, function(x,y) x+y)

outputVec
```

It is undoubtely more pratical to do this than it is to do the former way. This also allows you to be even more pratical:

```{r, eval=F}

outputVec <- innermap::innermap_dbl(inputVec, sum)
outputVec <- innermap::innermap_dbl(inputVec, `+`)
```

# Changing the distance between the elements used

In this case, we used each element with the following one, **but** we can also use it with different distances, using the argument `distance`. Supposing we want to multiply the `n` element with the `n+2` element we can do:

```{r}

outputVec2 <- innermap::innermap_dbl(inputVec, `*`, distance = 2)

outputVec2
```

The default distance between the elements is set to be one.

# A whole family

We used until now the `innermap_dbl` function, but in the same way we have `map`, `map_lgl`, `map_chr`,`map_int`, `map_raw`, `map_dfr`, `map_dfc` there is a `innermap`, `innermap_lgl`, `innermap_chr`,`innermap_int`, `innermap_dbl`, `innermap_raw`, `innermap_dfr`, `innermap_dfc`.


## Motivation

There is no native function in `purrr` that takes elements of the same vector or list and applies a function to them. As a researcher in the field of Structural Decomposition Analysis (SDA), I use a purrr::map quite often, and need quite often functions like the ones from `innermap` family. Since the Input-Output tables are quite heavy, there is no way one can give up the use of the optimized `purrr` functions. Also, the SDA methodology needs as input the original matrix, making functions like `accumulate` not suited for this.

I hope other will find this useful too! Really hope to receive comments about it! 

# Thanks

Big thanks to [pedroocava](https://github.com/pedrocava/) and [kimjoaoun](https://github.com/kimjoaoun/) for their early comments on this matter. Any mistakes are my own.

# Important to note

As I am an Economics PhD Student, I do not have the time, nor the skills, to make a hardcore low level coding in C++ to create super fast functions as is in the `purrr` package. Nonetheless, I still think this functions I presented perfom quite well, since I kept them simple and based in the `purrr` package. I already set an [issue](https://github.com/tidyverse/purrr/issues/797#issue-721645404) in the `purrr` Github repository today. If they add this feature in the `purrr` package, **what I hope they do**, I will not take this project forward or update it in anyway.
